Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-04 doc/paper.Rmd
changes to structure and information.Toolmark resolution length of markings included.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography \hh{sketch?}. Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted. We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

@@

>>>>

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography \hh{sketch?}. Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted. We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

@@

The land engraved markings across the cross section of the land are termed as bullet profiles \citep{aoas,ma2004}. These are often used for making land to land comparisons. Bullet signatures on the other hand \citep{chu2013,aoas} refer to a processed version of the raw land engraved markings or profiles. The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height. Loess fits are then used to model the structure. The residuals of these fits are called signatures. They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet. A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.

The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}. The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University. Both scans were generated using confocal microscopes but at different resolutions. The NIST database contains scans of the bullet sets Hamby 44 and 252 with the resolution 1.5625 $\mu m$ per pixel. The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.065 $\mu m$ per pixel. The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel. The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.

## Potential Challenges in Chumbley Score Adaptation

Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness. One way to investigate this is to evaluate bullet striations at different resolutions. This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm. Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces. Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets. Another aspect of investigation is the performance of the algorithm for different levels of smoothing when signatures are being compared as opposed to when raw markings or profiles are being compared.

The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion. They term them as windows of optimization and validation. The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.

The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length in mm, are smaller compared to toolmarks. Although in terms of digitzed pixels there is no straightforward way of establishing what window size is to be chosen. One important way to address this is to keep the number of windows of optimization sufficiently large or not too small. This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement. In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.

For too large window sizes, the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement is vastly reduced. On the other hand, this introduces another problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and all seem similar. The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all. This may lead to incorrect classification of bullets as coming from the same-source (or matches) or from different source (or non-matches).

##################

#### endedit
